Structured Versus Decorated Cospans

### 2020-12-04 cospans.tex
first take on conclusions

We have given conditions under which a decorated cospan double category is isomorphic to a structured cospan double category, in \cref{thm:equiv}. The converse question is also interesting: is every structured cospan double category equivalent to a decorated cospan double category? The answer is similar to the previous one: affirmative, under certain conditions that essentially allow us to make the passage from an appropriate functor $L\colon \A\to \X$ to an appropriate pseudofunctor $F\colon \A\to\Cat$, namely in the reverse direction than the earlier passage \cref{eq:FtoL}.

....

{\chris I don't know if the part below should be included or not, perhaps it is too speculative and not dead important.}

Even though for both \cref{thm:equiv} and the above described setting, the conditions required are only sufficient, they provide certain indication that under a suitable ``condition intersection'' in a sense, we could establish a functorial picture
\begin{displaymath}
\begin{tikzpicture}
\node (A) { \{Lax monoidal pseudofunctors $(\A,+)\to(\Cat,\times)$\} };
\node[below right=1.5 and 1.5 of A] (C)  { \{Symmetric monoidal double categories\} };
\node[below=1  of A] (D)  { \{Special opfibrations\} };
\node[below=2  of A] (E)  { \{Special laris\} };
\node[below=3 of A] (B)  { \{ Finitely cocomplete functors $A\to\X$\} };
\node[below=.05 of D] {$\cong$};
\path[->,font=\scriptsize]
(A) edge[bend left] node[above,sloped]{$F\mapsto F\lCsp$}  (C)
(D) edge (A)
(E) edge (B)
(B) edge[bend right] node[below,sloped]{$L\mapsto {}_{L}\lCsp$} (C);
\end{tikzpicture}
\end{displaymath}
with an actual natural isomorphism in the middle. This would imply that starting from an appropriate middle ground, employing the structured or the decorated cospan formalism is actually the same. We leave such considerations for future work.

>>>>

We have given conditions under which a decorated cospan double category is isomorphic to a structured cospan double category, in \cref{thm:equiv}. The converse question is also interesting: is every structured cospan double category equivalent to a decorated cospan double category? The answer is similar to the previous one: affirmative, under certain conditions that allow us to make the passage from an appropriate functor $L\colon \A\to \X$ to an appropriate pseudofunctor $F\colon \A\to\Cat$, namely in the reverse direction of the earlier \cref{eq:FtoL}.

Let us now give a sketch of the relevant story. Suppose $\A$ and $\X$ have finite colimits and $L\maps\A\to\X$ is a finite-colimit preserving functor -- which by \cref{thm:SC} are the required conditions for constructing the symmetric monoidal double category of structured cospans. If $L$ has a right adjoint `left inverse' (meaning the unit is the identity) $U\maps\X\to\A$, which moreover strictly preserves the choice of pushouts in $\A$ and $\X$, it can be shown\footnote{This will appear in the forthcoming paper \cite{CV}.} that $U$ is an opfibration. Consequently, $U$ corresponds to a pseudofunctor $F\colon \A\to\Cat$ by the reverse of the Grothendieck construction, i.e. the second part of \cref{thm:Grothendieck}. Furthermore, if $U$ preserves finite coproducts, by the special case of the cocartesian monoidal Grothendieck construction discussed under \cref{lem:MonGroth}, the corresponding pseudofunctor $F$ has a lax monoidal structure $(\A,+)\to(\Cat,\times)$. As a result, by \cref{DC}, $F$ has now enough structure to induce the symmetric monoidal double category of decorated cospans $F\lCsp$, and further it can then be verified that it is isomorphic to the one of structured cospans.

{\chris Give some sort of excuse why we don't do the full result here?? Because of missing - for us now - steps? Or other things?}

To give a better sense of how this lax monoidal pseudofunctor $F\maps\A\to\Cat$ is constructed, $F(a)$ is defined to be the pre-image of $a$ under $U$, namely the category of all objects in $x\in\X$ such that $U(x)=a$ and morphisms $k\maps x\to y\in\X$ such that $U(k)=1_a$. Given a morphism $f\maps a\to b$, there is a functor $F(f)\maps F(a)\to F(b)$ that maps some $x$ to the following pushout
\begin{displaymath}
 \begin{tikzcd}
La\ar[r,"Lf"]\ar[d,"\varepsilon_x"']\ar[dr,phantom,near end,"\ulcorner"] & Lb\ar[d] & \\
x\ar[d,dotted]\ar[r] & x+_{La}Lb\ar[d,dotted] & \textrm{in }\X \\
a\ar[r,"f"'] & b & \textrm{in }\A
 \end{tikzcd}
\end{displaymath}
where $\varepsilon_x\colon LU(x)=L(a)\to x$ is the counit of $L\dashv U$. The fact that $U$ strictly preserves pushouts is necessary to show that the pushout is mapped, via $U$, directly down to $b$.

{\chris I don't know if the part below should be included or not, perhaps it is too speculative and not dead important. 
%If we keep, I can add a bit more information about what these special opfibs and special laris are, if we think it helps -- or not if it's OK now. 
Bear in mind that me and Daniel are supposed to clarify the middle category isomorphism in this other paper too!}

Even though for both \cref{thm:equiv} and the above described setting, the conditions required are only sufficient, they provide certain indication that under a suitable ``condition intersection'' in a sense, we could establish a functorial picture
\begin{displaymath}
\begin{tikzpicture}
\node [align=center,draw] (A) { Lax monoidal pseudofunctors \\ $(\A,+)\to(\Cat,\times)$};
\node[below right=1.5 and 1.5 of A,align=center,draw] (C)  { Symmetric monoidal \\double categories };
\node[below left=1 and .5  of A,draw] (D)  {Special opfibrations};
\node[below left=2 and 1 of A,draw] (E)  {Special laris};
\node[below=3 of A,align=center,draw] (B)  {Finitely cocomplete \\ functors $A\to\X$};
\node[below=.01 of D] {\rotatebox{90}{$\cong$}};
\path[->,font=\scriptsize]
(A) edge node[above,sloped]{$F\mapsto F\lCsp$}  (C)
(D) edge (A)
(E) edge (B)
(B) edge node[below,sloped]{$L\mapsto {}_{L}\lCsp$} (C);
\end{tikzpicture}
\end{displaymath}
with an actual natural isomorphism in the middle. This would imply that starting from an appropriate middle ground, employing the structured or the decorated cospan formalism is essentially the same. We leave such considerations for future work.
